Contracts for Contracts

### 2023-06-26 paper/formal_model.tex
more formal section

The body of functions is simply a statement.
Statements can be a type-and-identifier declaration (without providing an initial
value), an expression (e.g. calling a function for its side effects), a
composition of two statements, an assignment, or a return statement.
The left-hand side part of assignments (i.e. assignable) can be either
a type-and-type declaration, an identifier, or a projection of mappings or structs.
Since \corelang permits both mappings and structs, which can recursively referring
other types, projection allows use to directly access nested data structures,
e.g. \code{v.field1[0].field2}.

Expressions could be an identifier, a literal value (e.g. numbers), a binary
operation, accessing mappings or structs, address calls and function calls.
An address call has form $e_1.f\Keywd{\{}(x: e_2)^*\Keywd{\}(}e_3^*\Keywd{)}$,
where $e_1$ is the target expression of type address,
$(x: e_2)^*$ are call options (e.g. specifying the gas budget),
and $e_3^*$ are call arguments.
An ordinary function call or struct constructor call is simpler, consisting
only of the function name and argument expressions.

@@

The syntax permits arbitrary expressions appeared as preconditions and
postconditions. There are additional well-typedness constraint.
The static semantics presented here concerns about when a specification is
well-typed.
Since \corelang is a subset of Solidity except for specifications, we omit the
static semantics for the rest of the language, referring to the informal
Solidity design document.

@@

what does it guarantee

>>>>

The body of functions is simply a statement.
Statements can be a type-and-identifier declaration (without providing an initial
value), an expression (e.g. calling a function for its side effects), a
composition of two statements, an assignment, a \Keywd{return} statement, or a \Keywd{revert} statement.
The left-hand side part of assignments (i.e. assignable) can be either
a type-and-type declaration, an identifier, or a projection of mappings or structs.
Since \corelang permits both mappings and structs, which can recursively referring
other types, projection allows use to directly access nested data structures,
e.g. \code{v.field1[0].field2}.
A \Keywd{revert} statement aborts the execution, reverting any changes made to
the blockchain state.

Expressions could be an identifier, a literal value (e.g. numbers), a binary
operation, projections of mappings or structs, address calls, and function calls.
An address call has form $e_1.f\Keywd{\{}(x: e_2)^*\Keywd{\}(}e_3^*\Keywd{)}$,
where $e_1$ is the target expression of type address,
$(x: e_2)^*$ are call options (e.g. specifying the gas budget),
and $e_3^*$ are call arguments.
An ordinary function call or struct constructor call is simpler, consisting
only of the function name and argument expressions.

@@

The syntax permits arbitrary expressions appeared as preconditions and
postconditions.
However, not all possible expression are valid executable specifications.
For example, a programmer can write \code{requires x + 1} as a precondition,
which do have a meaning for our purpose.
The static semantics presented here concerns about when a specification is
considered valid.

In general, the precondition and postcondition need to be well-typed, which
entails that the specified conditions are of type boolean, and only
use well-scoped variables.
Since \corelang is a subset of Solidity except for specifications, we focus on
the typing for specifications and omit the typing for the rest of the language,
referring to the informal Solidity design document.

@@

In this section, we discuss when and where \corelang provides the guarantee
that a violation will be detected (thus the execution of the smart contract
will be reverted).

Our translation semantics identifies the boundary
